### 2021-07-03 chapters/rq1/2_experiment_design.tex
minor tweaks exp2 design + results

\documentclass[../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we will design an experiment to answer our first research question.
We hypothesise that we can cluster mutants with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
\newline
We will start with extracting the characteristics from the selected mutation testing tool. 
Next we select and implement a clustering algorithm for which we can cluster mutants.
We can then cluster the mutants and compare the results of the clusters with the results of the mutants.

@@

\subsection{Validation}
The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the amount of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
The baseline for our experiment is the mutation score of the set of all mutants generated by our selected mutation testing tool.
\newline
To validate how effective our clustering is we can compare the weighed mutation score(see Chapter \ref{ch:weighed_score}) of our clustered set to the mutation score of the full set.
The closer the weighed score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants both types. 
We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
To decide if we accept or reject our hypothesis we use statistical significance\cite{Kirk1996PracticalCome}.
If the weighed mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.

\subsection{Mutant selection}
As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
To make our results reproducible we will select a random mutant based on a generated seed.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
Achieving consistent results while applying random selection will contribute to the validity of the experiment.
We will use the same statistical significance threshold to decide if the variation by the random sampling is acceptable or not.

>>>>

\documentclass[../../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we will design an experiment to answer our first research question.
We hypothesise that we can cluster mutants with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
\newline
We will start with extracting the characteristics from the selected mutation testing tool. 
Next we select and implement a clustering algorithm for which we can cluster mutants.
We can then cluster the mutants and compare the results of the clusters with the results of the mutants.

@@

\subsection{Validation}
\label{ch:exp1_validation}
The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the amount of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
The baseline for our experiment is the mutation score of the set of all mutants generated by our selected mutation testing tool.
\newline
To validate how effective our clustering is we can compare the weighed mutation score(see Chapter \ref{ch:weighed_score}) of our clustered set to the mutation score of the full set.
The closer the weighed score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants both types. 
We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
If the weighed mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.

\subsection{Mutant selection}
As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
To make our results reproducible we will select a random mutant based on a generated seed.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
We will repeat our experiment 30 times with 30 different seeds.
Achieving consistent results while applying random selection will contribute to the validity of the experiment.
We will use the same statistical significance threshold to decide if the variation by the random sampling is acceptable or not.

### 2021-07-03 chapters/rq2/1_experiment_design.tex
minor tweaks exp2 design + results

\documentclass[../../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we will design an experiment to answer our second research question.
We hypothesise that we can train a machine learning model with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
\newline
\todo{finish this}

\subsection{Selecting a machine learning model}
The first choice we have to make on selecting a machine learning model is to decide on supervised or unsupervised learning.
The application and description of these two types of learning are described in Chapter \ref{ch:background}.
Supervised machine learning models require a labeled data set.
Our data set consists of mutants represented by the characteristics identified in Chapter \ref{ch:clustering_characteristics}.
This data set is unlabeled as it does not map the mutants to categories.
We also do now know what the categories would be or how to map a mutant to a specific category.
Instead of manually evaluating each mutant and identifying categories to map them to we can train a machine learning model to do this for us.
To do this we need an unsupervised machine learning model.
\newline
Our starting point for selecting an unsupervised machine learning model are the models discussed in the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
We will select a model by a process of elimination.

\subsubsection{Mountain model}
The mountain clustering model is an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
While finding out the number of clusters we need might be valuable it is not what we want to achieve. 
Therefore we will not select mountain clustering.

@@

\subsubsection{Self organizing map, C-means and fuzzy C-means}
The self organizing map, C-means and fuzzy C-means models are all models that would fit our purpose. 
To choose between these models we will look at the performance of each model.
Mingoti et al., compared the performance of these three models \cite{Mingoti2006ComparingAlgorithms}.
They did a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation\cite{Mingoti2006ComparingAlgorithms}.
Other variables such as cluster boundaries and overlap were also controlled in the experiment.
They observe that all the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
The best results for average recovery and internal dispersion rates were found for Fuzzy C-means which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
The C-means method was very affected by the presence of a large amount of outliers.
They conclude that the self organizing map did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
They also conclude that their results partially overlap with other research in the same domain.
\newline
Based on these results we can conclude that the fuzzy c-means model has the best performance overall.
For this experiment we select the fuzzy c-means model.

\subsection{Training strategy}
Our data set contains of mutants represented by their characteristics. 
The characteristics of the mutants are generated per project and have a project specific context.
The relations between the different mutants are also specific per project.
For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
To keep the relations and context we decided to train our model per project.
\newline
To validate our model we need to split our data set into a training set and a validation set.
Our total data set consists of approximately 1 million mutants generated over fourteen projects.
Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
We will divide our projects into training and validation groups to reflect the 80/20 ratio.

\subsection{Implementation of algorithm}
The origin of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
The fuzzy c-means algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
Based on their research an implementation can be developed.
There are existing libraries that contain the implementation of the fuzzy c-means algorithm.
Developing our own version of the fuzzy c-means algorithm is out of scope for this research.
For this reason we choose to use a library that implements the fuzzy c-means algorithm.
The library we will use for this experiment is the fuzzy c-means implementation build by Dias et al.\cite{dias2019fuzzy}.
The theory developed by Bezdek et al., is implemented in this library.
The library has been used in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020} and is actively maintained.

\subsection{Algorithm parameters}

\subsection{Validation}

\end{document}

>>>>

\documentclass[../../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we will design an experiment to answer our second research question.
We hypothesise that we can train a machine learning model with the set of characteristics we identified
while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
We will select a machine learning model, decide on a training strategy, train the model and validate the results.

\subsection{Selecting a machine learning model}
We will first choose between supervised or unsupervised learning.
Our starting point for selecting a machine learning model are the models discussed in the survey done by K.L. Du.
Based on the choice of supervised or unsupervised learning we will review the corresponding models and decide if they fit our use case by a process of elimination.

\subsubsection{Supervised or unsupervised learning}
In this subsection we will decide on supervised or unsupervised learning.
The application and description of these two types of learning are described in Chapter \ref{ch:background}.
\newline
Supervised machine learning models require a labeled data set.
We can argue that we can label our data set with the results of our first experiments.
Each mutant would then be labeled with the identity of the cluster it is clustered in by the hierarchical clustering model.
If we would train a supervised model with the data labeled as described, it may result in a model that can do the same as hierarchical clustering.
Our research is not to prove if we can train a model to cluster mutants the same way as hierarchical clustering.
We want to train a model to find the relations between the characteristics without having to define these relations manually.
\newline
Unsupervised learning models do not need a labeled data set and can recognize relations or discover hidden patterns in the characteristics we identified.
This corresponds to what we want to achieve with this experiment.
Therefore we choose unsupervised learning to continue with in our experiment.

\subsubsection{Mountain and subtractive clustering}
The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
\newline
The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points. Each grid point is a potential cluster center. The potential for each grid is calculated based on the density of the surrounding data points. The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
\newline
Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
To find out the number of clusters we need for our model we can make use of these models.

@@

\subsubsection{Self organizing map, C-means and fuzzy C-means}
The self organizing map, C-means and fuzzy C-means models are all models that would fit our purpose. 
To choose between these models we will look at the performance of each model.
Mingoti et al., compared the performance of these three models \cite{Mingoti2006ComparingAlgorithms}.
They did a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation\cite{Mingoti2006ComparingAlgorithms}.
Other variables such as cluster boundaries and overlap were also controlled in the experiment.
They observe that the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
The best results for average recovery and internal dispersion rates were found for Fuzzy C-means which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
The C-means method was very affected by the presence of a large amount of outliers.
They conclude that the self organizing map did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
They also conclude that their results partially overlap with other research in the same domain.
\newline
Based on these results we can conclude that the fuzzy c-means model has the best performance overall.
For this experiment we select the fuzzy c-means model.

\subsection{Training strategy}
\label{ch:training_strategy}
Our data set contains of mutants represented by their characteristics. 
The characteristics of the mutants are generated per project and have a project specific context.
The relations between the different mutants are also specific per project.
For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
To keep the relations and context we decided to train a model per project.
\newline
To validate the model we need to split our data set into a training set and a validation set.
Our total data set consists of approximately 1 million mutants generated over fourteen projects.
Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
We will divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.

\subsection{Implementation of algorithm}
The research paper of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
The fuzzy c-means algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
Based on their research an implementation can be developed.
There are existing libraries that contain the implementation of the fuzzy c-means algorithm.
Developing our own version of the fuzzy c-means algorithm is out of scope for this research.
For this reason we choose to use a library that implements the fuzzy c-means algorithm.
The library we will use for this experiment is the fuzzy c-means implementation build by Dias et al.\cite{dias2019fuzzy}.
The theory developed by Bezdek et al., is implemented in this library.
The library has been used in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020} and is actively maintained.

\subsection{Algorithm parameters}
The fuzzy c-means algorithm has to initialize the centroids of the clusters.
Our selected library does this based on a random number.
It also allows us to control this random number.
To make the initialization of the centroid reproducible and deterministic we will use a generated seed.
The seed generated is included in the Github repository\cite{rbasarat-repo}.
\newline
Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995}.
They researched what effect which parameters have on the validity of the clusters.
They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
The value of \textit{m} in this experiment is 2.
\newline
Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
Fortunately there are some techniques available to estimate the cluster amount.
During model selection we discussed the mountain and subtractive models. 
These models are designed to estimate the amount of clusters in a data set.
The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
We will select and implement the subtractive clustering model to estimate the amount of clusters.
We can use this amount for the cluster amount parameter in the fuzzy c-means algorithm.

\subsection{Implementation of subtractive clustering}
The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
Based on the paper of the theory of the subtractive algorithm\cite{Chiu1994FuzzyEstimation} we found an existing implementation\cite{matlabSubtractive}.
The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.

\subsection{Validation}
To validate our machine learning model we will calculate the weighed mutation score and cluster accuracy based on the clusters calculated by the model.
This is the same validation method as described in Chapter \ref{ch:exp1_validation}.
We will use the mutants of projects that are not selected for training as input for validating the model.
As described in Chapter \ref{ch:training_strategy} the projects should represent approximately 20\% of the total amount of mutants in our data set.
It may be possible the model has to cluster mutants it has never seen before.
We will use the same statistical significance threshold as in our firsts experiment to decide whether we will accept or reject or hypothesis.
To make our results reproducible we will select a random mutant based on a generated seed.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
\end{document}
